Horse Racing Analysis

### 2023-12-15 analysis/new_analysis.qmd
ready for merge

---
title: "Horse Racing Analysis"
author: "MC"
format:
  html:
    theme: flatly
    date: last-modified
    smooth-scroll: true
    cap-location: top
    code-fold: true
execute:
  warning: false
---

@@

```{r}
#| label: setup
library(tidyverse)
library(lme4)
library(here)

@@

Now draw a graph.  NB., we have two options here:  Either we scale up all of the bets so that they're "out of 100" or we leave as-is.  Starting with as-is:

```{r}
df_ready |> ggplot(aes(x=delivery,y=BET100,fill=speaker)) +
  geom_bar(position="dodge",stat="summary",fun.y="mean")
```

>>>>

---
title: "Horse Racing Analysis"
author: "MC"
format:
  html:
    theme: flatly
    date: last-modified
    smooth-scroll: true
    cap-location: top
    code-fold: true
    embed-resources: true
execute:
  warning: false
---

@@

```{r}
#| label: setup
library(tidyverse)
#library(lme4)
library(buildmer)
library(here)

@@

So I'm as sure as I can be that these data are the same (or at least created to the same standards) as Espe's.  Now draw a graph.  NB., we have two options here:  Either we scale up all of the bets so that they're "out of 100" or we leave as-is.  As-is is on the left, scaled bet is on the right.

```{r}
#| layout-ncol: 2

df_plot <- df_ready |> group_by(delivery,speaker) |> summarise(mb=mean_se(BET),mb100=mean_se(BET100))

df_plot |> ggplot(aes(x=delivery,y=mb$y,fill=speaker)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=mb$ymin,ymax=mb$ymax),width=.4,position=position_dodge(.9))
df_plot |> ggplot(aes(x=delivery,y=mb100$y,fill=speaker)) +
  geom_bar(position="dodge",stat="identity") +
  geom_errorbar(aes(ymin=mb100$ymin,ymax=mb100$ymax),width=.4,position=position_dodge(.9))
```

Perhaps violin plots might make the distributions a bit clearer:

```{r}
#| layout-ncol: 2
df_ready |> ggplot(aes(x=delivery,y=BET,fill=speaker)) +
  geom_violin(alpha=.5) +
  geom_boxplot(width=.2,position=position_dodge(.9)) 
df_ready |> ggplot(aes(x=delivery,y=BET100,fill=speaker)) +
    geom_violin(alpha=.5) +
    geom_boxplot(width=.2,position=position_dodge(.9))
```

### Analysis

Let's fit the maximal possible model to the data:

```{r}
#| output: false
modB <- buildmer(
  BET ~ delivery * speaker +
    (delivery * speaker | ppt) +
    (delivery * speaker | horse),
  buildmerControl=buildmerControl(direction='order',
    args=list(control=lme4::lmerControl(optimizer='bobyqa'))),
  data = df_ready
)
```
OK, so the maximal converging model is **`r deparse1(formula(modB@model))`**.  Coefficents are shown below.

```{r}
library(gtsummary)
tbl_regression(modB@model,intercept=TRUE)
```

Now let's repeat the analysis, using the DV in which all bets are scaled to a total of 100.

```{r}
#| output: false
modB1 <- buildmer(
  BET100 ~ delivery * speaker +
    (delivery * speaker | ppt) +
    (delivery * speaker | horse),
  buildmerControl=buildmerControl(direction='order',
    args=list(control=lme4::lmerControl(optimizer='bobyqa'))),
  data = df_ready
)
```

The maximal converging model is **`r deparse1(formula(modB1@model))`**. Coefficients:

```{r}
tbl_regression(modB1@model,intercept=TRUE)
```

---

That's as far as I got, but it seems to confirm Espe's original findings.  Perhaps the violin plots above give some clues as to why there is no interaction?
